Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-07-02 09_Article/Simon_2021_Bone_Draft_V02.tex
Article corrections

Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels. Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well. Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls. Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different. At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

High resolution peripheral quantitative computed tomography (HR-pQCT) scans allow one to perform \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) in the distal radius and distal tibia \cite{Boutroy2005}. Moreover, the image obtained from HR-pQCT can be used for finite element analysis (FEA) to predict mechanical properties \cite{Boutroy2008}. Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from the HR-pQCT scan that can be used to assess bone strength within a reasonable computation time \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,AriasMoreno2019}. Thus, it could be legitimate to use hFE for OI patient's bone strength estimation and potentially fracture risk assessment. However, HR-pQCT-based FEA relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to investigate the hypothesis of similar fabric-elasticity relationships.

>>>>

Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels. Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well. Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls. Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different. At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior (ultimate stress and ultimate strain) than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}. Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,AriasMoreno2019}. Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to investigate the hypothesis of similar fabric-elasticity relationships.
